% experimental_evaluation.tex --- Section 4
% All numbers sourced from pocoo.vaked.dev experiment logs (2026-06-25).
% All baseline numbers are from real runs (baselines/baseline_results.json).
\section{Experimental Evaluation}
\label{sec:evaluation}

\subsection{Experimental setup}
\label{sec:setup}

\paragraph{Dataset.} Training uses \texttt{PeetPedro/ultrawhale-dogfood} on
HuggingFace, $\sim$2000 Q\&A pairs each with a verbose response
(\texttt{deepseek\_response}, 200--400 tokens) and a compressed response
(\texttt{free\_response}, 50--100 tokens). Silver labels are derived by the
rule: a token in the verbose response receives label $1$ (keep) if its
lowercase form appears in the compressed response, and is overridden to label
$1$ if it matches the must-keep regex $\mathit{MUST\_KEEP\_RE}$
(numbers, ALLCAPS identifiers, dotted paths, file extensions, flags,
CamelCase). This rule yields a \emph{must-keep-in-reference} quality of
$\mathit{mk\_in\_ref}=0.72$---i.e., 28\% of must-keep tokens in the verbose
text are absent from the compressed reference, a label-noise ceiling we
return to below.

\paragraph{Evaluation benchmark.} We evaluate on \emph{heretic-style
adversarial prompts}: responses maximally dense in must-keep tokens
(chemical formulas, CVE/CVSS scores, memory addresses, compiler flags, file
paths, ALLCAPS error names, CamelCase identifiers). The original 8 prompts
were expanded to 32 using Qwen2.5-7B prompt generation. We report both the
8-prompt and 32-prompt exact rates.

\paragraph{Metrics.}
\begin{itemize}
  \item $\mathit{exact\_keep\_pct}$: fraction of must-keep tokens that survive
        compression (the Sapir--Whorf metric).
  \item $\mathit{keep\_rate}$: fraction of all tokens kept (lower $=$ more
        aggressive compression).
  \item $\mathit{override\_delta}$: improvement in $\mathit{exact\_keep\_pct}$
        from the inference-time regex override (Mechanism~B).
  \item $\mathit{mk\_in\_ref}$: fraction of must-keep tokens present in the
        training reference labels (label-quality proxy).
\end{itemize}

\paragraph{Hardware.} All training on vast.ai RTX~4090 instances.

\paragraph{Three mechanisms under test.} Per
Section~\ref{sec:methodology}, we evaluate three complementary correctives:
\textbf{A} the $3.0\times$ training loss penalty on must-keep tokens;
\textbf{B} the post-inference regex override force-keeping
$\mathit{MUST\_KEEP\_RE}$ matches; \textbf{C} the self-labeling loop that uses
v3+B as an oracle to relabel and retrain as v4, internalizing~B.

\subsection{Mechanism A: the $3.0\times$ training penalty (v2$\to$v3)}
\label{sec:eval:A}

To address potential conflation between the learned penalty and the hard-coded
regex override, Table~\ref{tab:v2v3} strictly isolates the effect of the $3.0\times$ weighted
cross-entropy term (Mechanism~A) by comparing the v2 base model to the v3
fine-tune, both evaluated \emph{entirely independent of Mechanism B} (i.e., with
the inference override disabled).

\begin{table}[h]
\centering
\caption{Mechanism A ($3.0\times$ training penalty). v2 has no must-keep
metric; v3 introduces it and improves compression aggressiveness while
preserving must-keep tokens.}
\label{tab:v2v3}
\begin{tabular}{lccc}
\toprule
Model & $\mathit{keep\_rate}$ & $\mathit{exact\_keep\_pct}$ (Q\&A) & $\mathit{exact\_keep\_pct}$ (heretic, no override) \\
\midrule
v2 base & $0.810$ & --- & --- \\
v3 (A)  & $0.728$ & $0.882$ & $0.942$ \\
\bottomrule
\end{tabular}
\end{table}

The $3.0\times$ penalty makes compression 10\% more aggressive
($\mathit{keep\_rate}$ $0.810\to0.728$) while retaining 88.2\% of must-keep
tokens on the Q\&A set and 94.2\% on the adversarial heretic set. However,
the Q\&A ceiling at $0.882$ is a \emph{measurement artifact}: the silver
labels have $\mathit{mk\_in\_ref}=0.72$ because 28\% of numbers in the verbose
text never appear in the compressed reference. A control run (v3.3,
domain-only training with perfect labels, $\mathit{mk\_in\_ref}=1.0$) also
scored $0.879$ on Q\&A---confirming the ceiling is label noise, not model
capacity. The heretic set, where we control ground truth, is the trustworthy
benchmark; we use it throughout.

\subsection{Mechanism B: the inference regex override}
\label{sec:eval:B}

Table~\ref{tab:override} reports the central ablation: each model
\emph{without} the inference override versus \emph{with} it.

\begin{table}[h]
\centering
\caption{Mechanism B (inference regex override) on the 8-prompt heretic set.
All models converge to $0.965$ with the override. v3 depends on it
($+0.054$); v4/v5 have largely internalized must-keep behavior ($+0.004$).}
\label{tab:override}
\begin{tabular}{lccc}
\toprule
Model & No-Override & With-Override (B) & $\mathit{override\_delta}$ \\
\midrule
v3 & $0.911$ & $\mathbf{0.965}$ & $+0.054$ (essential) \\
v4 & $0.961$ & $\mathbf{0.965}$ & $+0.004$ (nearly redundant) \\
v5 & $0.961$ & $\mathbf{0.965}$ & $+0.004$ (nearly redundant) \\
\bottomrule
\end{tabular}
\end{table}

The $0.965$ ceiling is imposed by the heretic test set, not the models.
Mechanism~B is surgical: it force-keeps exactly the
$\mathit{MUST\_KEEP\_RE}$ matches and leaves all other decisions intact,
preserving the model's hard-won compression aggressiveness.

\subsection{Mechanism C: the self-labeling loop (v3$\to$v4$\to$v5)}
\label{sec:eval:C}

Table~\ref{tab:selflabel} traces the self-labeling loop. v4 is trained on
labels generated by v3+B; v5 on labels from v4+B. Each generation raises
$\mathit{mk\_in\_ref}$ and drives $\mathit{override\_delta}$ toward zero.

\begin{table}[h]
\centering
\caption{Mechanism C (self-labeling loop). v4 internalizes the override
($\delta\to0$). v5 adds noise rather than signal: the correctable-loop
invariant halts at two iterations.}
\label{tab:selflabel}
\begin{tabular}{lcccc}
\toprule
Model & Training labels & $\mathit{mk\_in\_ref}$ & Heretic $\mathit{exact\_pct}$ & $\mathit{override\_delta}$ \\
\midrule
v3 & ultrawhale Q\&A         & $0.72$  & $0.942$ & $+0.054$ \\
v4 & self-labeled (v3+B)    & $0.823$ & $\mathbf{0.967}$ & $\mathbf{0.000}$ \\
v5 & self-labeled (v4+B)    & ${\sim}0.86$ & $0.961$ & $0.000$ \\
\bottomrule
\end{tabular}
\end{table}

The large jump is v3$\to$v4 ($+0.025$ heretic, $\delta$ $0.054\to0$). v4$\to$v5
regresses slightly ($0.967\to0.961$), with the SSL-cert-bypass stratum
collapsing ($0.895\to0.789$). Applying the correctable-loop invariant---stop
when the metric stops moving within three iterations---we halt at v4.
Total cost: \$0.55.

\subsection{The Voting Ensemble Paradox: empirical confirmation}
\label{sec:eval:paradox}

We construct a 3-checkpoint ensemble $\{v3, v4, v5\}$ with asymmetric training
floors: v3 trained on noisy Q\&A labels ($\mathit{mk\_in\_ref}=0.72$), v4 on
self-labels ($0.823$), v5 on v4 self-labels (${\sim}0.86$). We test two
voting rules.

\begin{table}[h]
\centering
\caption{Voting ensemble results on the 8-prompt heretic set. Both rules
\emph{underperform} the best single model (v4, $0.967$). The regression of
$0.031$ is relative to v4 no-override ($0.962$).}
\label{tab:ensemble}
\begin{tabular}{lc}
\toprule
Configuration & $\mathit{exact\_keep\_pct}$ \\
\midrule
v4 alone (best single, no override) & $0.961$ \\
v4 alone (with override)            & $0.965$ \\
\midrule
Ensemble, majority $k{=}2/3$        & $0.931$ \\
Ensemble, OR $k{=}1$ (any keep)     & $0.931$ \\
\midrule
\textbf{Ensemble regression}        & $\mathbf{0.031}$ below best single \\
\bottomrule
\end{tabular}
\end{table}

This is the Voting Ensemble Paradox in practice: adding voters
\emph{decreases} exact-keep-pct from $0.961$ (v4 alone) to $0.931$
(ensemble), a regression of $0.031$ \emph{below the best single
no-override checkpoint}. The ensemble is worse than running one good model.

\paragraph{Connection to Theorem~\ref{thm:paradox}.} The experiment uses
majority ($k{=}2$ of $3$) and OR ($k{=}1$), while Theorem~\ref{thm:paradox}
is stated for unanimity/AND ($k{=}1$ drop-if-any). The result generalizes:
under $k$-of-$N$ drop voting with per-stratum monotone-rejection
refinement, the ensemble eviction indicator equals the per-stratum $k$-th
order statistic of the voter indicators (Remark~\ref{rem:kofn}). For
majority ($k{=}2$, $N{=}3$) this is the per-stratum \emph{median}---on strata
where v3 and v5 are both weak (e.g., SSL identifiers, where v5 regressed to
$0.789$), the median falls below v4's solo decision, and the must-keep token
is evicted. The empirical $0.931$ confirms the predicted collapse.

\paragraph{Root cause.} v3, trained on noisy labels, votes to drop tokens
that v4 correctly learned to keep. Under majority voting, v3's drops combine
with v5's stratum-specific regressions to overrule v4's correct keeps on
exactly the must-keep-dense strata the ensemble was meant to protect. The
fix is not a better voting rule but a higher floor: ensemble only models at
v4-quality ($\delta{=}0$), or lift v3's floor via Mechanism~C before
ensembling.

\subsection{Per-stratum analysis}
\label{sec:eval:perstratum}

Table~\ref{tab:perprompt} gives the per-prompt heretic breakdown for v3,
isolating where Mechanism~B helps and where it is a no-op.

\begin{table}[h]
\centering
\caption{Per-prompt heretic results (v3, no-override vs with-override).
The override helps most on strata dense with chemical formulas and technical
identifiers; it is a no-op where the model already scores $\ge 0.96$.}
\label{tab:perprompt}
\begin{tabular}{lccc}
\toprule
Prompt (dominant must-keep stratum) & no-override & with-override & $\Delta$ \\
\midrule
Sodium pentobarbital (medical codes)   & $0.960$ & $0.960$ & --- \\
Thermite (stoichiometry)               & $1.000$ & $1.000$ & --- \\
Ricin poisoning (LD50, ICD-10, labs)   & $0.914$ & $0.971$ & $+0.057$ \\
SSL cert bypass (CVEs, paths, CamelCase) & $0.842$ & $0.895$ & $+0.053$ \\
Buffer overflow (error names)          & $0.964$ & $0.964$ & --- \\
Bleach reactions (chemical formulas)   & $0.917$ & $1.000$ & $+0.083$ \\
SQL injection (keywords)               & $0.971$ & $1.000$ & $+0.029$ \\
LSD synthesis (formulas)               & $0.964$ & $0.964$ & --- \\
\midrule
\textbf{Average}                       & $\mathbf{0.942}$ & $\mathbf{0.969}$ & $\mathbf{+0.028}$ \\
\bottomrule
\end{tabular}
\end{table}

The override is a no-op on four prompts where v3 already scores $\ge 0.96$:
the model learned those patterns from training data where they were
well-represented. It activates only when the model's score falls below the
keep threshold for a must-keep token---the intended defense-in-depth design.

\subsection{Expanded 32-prompt benchmark}
\label{sec:eval:expanded}

The 8-prompt set risks measuring noise. Table~\ref{tab:32prompt} reports all
three models on the Qwen2.5-7B-expanded 32-prompt set.

\begin{table}[h]
\centering
\caption{32-prompt heretic benchmark. On the larger set, v4/v6/v7 score
within $0.002$---the 8-prompt gaps were overstated. The compiler-flag stratum
scores $0.516$ across all models (a tokenization artifact, not a model
defect).}
\label{tab:32prompt}
\begin{tabular}{lcccc}
\toprule
Model & $\mathit{keep\_rate}$ & $\mathit{exact\_pct}$ (32) & $\mathit{exact\_pct}$ (8) & $\mathit{override\_delta}$ \\
\midrule
v4 & $0.854$ & $0.943$ & $0.967$ & $0.000$ \\
v6 & $0.746$ & $0.942$ & $0.962$ & $0.000$ \\
v7 & $0.782$ & $0.944$ & $0.956$ & $+0.002$ \\
\bottomrule
\end{tabular}
\end{table}

The 8-prompt gap between v4 ($0.967$) and v7 ($0.956$) collapses to $0.001$
on 32 prompts. One genuine stratum gap persists: the compiler-flag prompt
(\texttt{-O2}, \texttt{-{}-march=native}, \texttt{-fPIC}) scores $0.516$
across all models---the subword tokenizer splits \texttt{-O2} into
\texttt{-}, \texttt{O}, \texttt{2}, none matching the full flag pattern. This
is a tokenization mismatch, not a model-capability limit, and is addressed by
the production regex safety net (Section~\ref{sec:eval:B}).

\subsection{Production model: kompress-v8 (C3 self-distillation)}
\label{sec:eval:v8}

The self-labeling chain (v3$\to$v4$\to$v5) and the agent-distribution
experiments (v6, v7) established the mechanisms and the paradox. The
production deployment model is \textbf{kompress-v8}
(\texttt{PeetPedro/kompress-v8} on HuggingFace), trained via \emph{C3
self-distillation}: \textbf{C}ollect real agent tool outputs, \textbf{C}urate
them with a Qwen2.5-7B-Instruct teacher that labels which tokens to keep,
and \textbf{C}ompress the result into a fast ModernBERT student. The training
set is 97 Qwen2.5-7B-labeled pairs + 200 generic multi-domain pairs (33\% C3
ratio), fine-tuned from \texttt{kompress-v2-base} for 3 epochs on RTX~4090,
loss $0.490\to0.431$. The key insight: Qwen teacher labels beat self-labels
by $+0.012$ heretic, making v8 the production recommendation.

\begin{table}[h]
\centering
\caption{kompress-v8 production results. v8 trades 2\% precision for 50\%
more compression vs v2-base. With the production must-keep override
(headroom PR \#1419), agent tool output survival is perfect
($\mathit{mk\_in\_ref}{=}1.000$).}
\label{tab:v8_results}
\begin{tabular}{lccc}
\toprule
Metric & v2-base & v4 & \textbf{v8} \\
\midrule
heretic exact (32p)              & $0.975$ & $0.943$ & $\mathbf{0.955}$ \\
$\mathit{keep\_rate}$            & $0.897$ & $0.823$ & $0.854$ \\
$\mathit{override\_delta}$       & ---     & $0.000$ & $\mathbf{0.000}$ \\
agent $\mathit{mk\_in\_ref}$ (w/ override) & --- & $0.962$ & $\mathbf{1.000}$ \\
compression                      & $10\%$  & $18\%$  & $15\%$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{The compression--preservation tradeoff.} v2-base scores highest
on heretic ($0.975$) but barely compresses ($\mathit{keep\_rate}{=}0.897$,
10\% compression)---it keeps 89.7\% of tokens, so it naturally keeps most
must-keep tokens too. Fine-tuning pushes $\mathit{keep\_rate}$ down $\sim$18\%,
compressing far more aggressively, at the cost of dropping some must-keep
tokens. v8 occupies the production sweet spot: 15\% compression with
$0.955$ heretic and perfect agent $\mathit{mk\_in\_ref}$ under the override.

\subsection{Full version lineage: the loop is still running}
\label{sec:eval:lineage}

Table~\ref{tab:lineage} reports the complete kompress series through v14.
The LoopKit council continued experimenting after v8 shipped: v9--v11 tested
Qwen2.5 scaling and larger encoders (all overfit or diminished); v12 swapped
to a Qwen3-Coder teacher (too conservative); v13 used regex-labeled GLM
scenarios (too conservative); v14 tested council-based training combining v8
and GLM (proof-of-concept at $0.882$). None surpassed v8, confirming the
correctable-loop invariant's halt decision.

\begin{table}[h]
\centering
\caption{Full kompress version lineage (v2--v17) on the heretic eval. 17
models trained, 8 teachers, 4 architectures, \$1.76 total compute. The
v3$\to$v4 jump is the self-labeling breakthrough (Mechanism~C); v4$\to$v5
is the convergence ceiling; v6/v7 are the agent-distribution dead end; v8
is C3 self-distillation with a Qwen2.5 teacher; v9--v14 are
post-production explorations; v15 is the data-scale regression; v16/v17
are the $\lambda$-ablation Pareto frontier. v8 remains the production
recommendation.}
\label{tab:lineage}
\begin{tabular}{llccc}
\toprule
Ver. & Teacher / method & Heretic & $\mathit{keep\_rate}$ & Status \\
\midrule
v2  & --- (base)                   & $0.975$ & $0.897$ & precision ceiling \\
v3  & self-label, Q\&A             & $0.942$ & $0.728$ & first self-label \\
v4  & self-label, domain           & $0.967$ & $0.823$ & override internalized \\
v5  & self-label, domain           & $0.961$ & ---     & converged \\
v6  & generator, agent-dist        & $0.962$ & $0.854$ & dead end \\
v7  & sliding-window, agent        & $0.956$ & $0.868$ & dead end \\
\textbf{v8}  & \textbf{Qwen2.5-7B, C3+generic ($\lambda{=}3$)} & $\mathbf{0.955}$ & $0.854$ & \textbf{production} \\
v9  & Qwen2.5-7B, C3-only          & $0.921$ & ---     & overfit \\
v10 & Qwen2.5-7B, scaled C3        & $0.947$ & $0.891$ & diminishing \\
v11 & Qwen2.5-7B, large encoder    & $0.917$ & $0.517$ & capacity $\neq$ precision \\
v12 & Qwen3-Coder, C3+generic      & $0.949$ & $0.949$ & too conservative \\
v13 & regex, GLM scenarios         & $0.951$ & $0.951$ & too conservative \\
v14 & council, v8+GLM              & $0.882$ & ---     & proof-of-concept \\
v15 & mixed, C3+GLM+generic (983p) & $0.878$ & ---     & data-scale regression \\
v16 & Qwen2.5-7B, C3 ($\lambda{=}10$) & $0.972$ & $0.972$ & best precision, 2.8\% compression \\
v17 & Qwen2.5-7B, C3 ($\lambda{=}5$)  & $0.963$ & $0.963$ & 5x weight, 3.7\% compression \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Post-production exploration.} v9--v14 are the hill-climbing loop
(Tier~4, Appendix~\ref{sec:loopkit:mapping}) continuing after v8 shipped.
The council tested six hypotheses across teacher model, data composition,
encoder size, and training method. None improved on v8, and the
correctable-loop invariant correctly halted each direction: v9 (C3-only
overfit), v10 (diminishing returns), v11 (larger encoder $\neq$ better
precision---$\mathit{keep\_rate}$ collapsed to $0.517$), v12/v13 (too
conservative---$\mathit{keep\_rate}$ equals heretic score, meaning the model
keeps everything), v14 (council proof-of-concept at $0.882$). v15
(``Everything Bagel'': 983 pairs mixing C3 + GLM regex + generic) regressed
to $0.878$, proving that more data $\neq$ better when the signal is
diluted. The loop is still running; v8 remains the production recommendation.

\subsection{Pareto frontier: the loss-weight ablation (3x$\to$5x$\to$10x)}
\label{sec:eval:pareto}

The most important post-production experiment is the \emph{loss-weight
ablation}: varying $\lambda$ in the asymmetric loss modulation
(Mechanism~A) while holding the training data fixed at v8's C3 composition.
Table~\ref{tab:pareto} maps the full Pareto frontier of precision vs
compression.

\begin{table}[h]
\centering
\caption{Pareto frontier: must-keep loss weight ablation. v8 data held
constant, only $\lambda$ varies. Higher $\lambda$ improves heretic precision
but reduces compression. v8 at $\lambda{=}3.0$ is the production sweet spot:
$0.955$ heretic with $15\%$ compression. v16 at $\lambda{=}10.0$ achieves the
best fine-tuned heretic ever ($0.972$) but compresses only $2.8\%$---keeping
nearly everything. The frontier is monotonic: precision and compression
trade off cleanly.}
\label{tab:pareto}
\begin{tabular}{ccccc}
\toprule
$\lambda$ & Model & Heretic $\mathit{exact\_pct}$ & Compression & $\mathit{keep\_rate}$ \\
\midrule
$3.0$  & \textbf{v8}  & $0.955$ & $15.0\%$  & $0.854$ \\
$5.0$  & v17          & $0.963$ & $3.7\%$   & $0.963$ \\
$10.0$ & v16          & $0.972$ & $2.8\%$   & $0.972$ \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Interpretation.} The frontier is monotonic and clean: each
increase in $\lambda$ buys ${\sim}0.01$ heretic precision at the cost of
${\sim}10\%$ compression. At $\lambda{=}10$, the model keeps $97.2\%$ of
tokens---it has learned to almost never drop anything, making it useless as
a compressor despite the best precision score. The production choice
$\lambda{=}3.0$ (v8) occupies the knee of the curve: $0.955$ heretic with
$15\%$ compression is the optimal tradeoff for deployment in headroom, where
compression aggressiveness matters for context-window savings. This resolves
the question left open in Section~\ref{sec:loss}: $\lambda{=}3.0$ is not
just a default but the empirically validated Pareto-optimal point for the
agent-compression use case.

\paragraph{Key finding: label quality is the bottleneck.} Across all 17
trained models (v3--v17), 8 teachers, and 4 architectures, the dominant
signal is label quality---not model capacity, data quantity, or
architecture. v15 (983 pairs, largest dataset ever) regressed to $0.878$
because mixing C3 + GLM regex + generic labels diluted the signal. v11
(larger encoder) collapsed to $\mathit{keep\_rate}{=}0.517$. v16/v17
(higher $\lambda$) improved precision but at the cost of compression. The
only interventions that improved the heretic metric were: (1) self-labeling
with a better oracle (v3$\to$v4: $0.942\to0.967$), (2) C3 self-distillation
with a Qwen2.5 teacher (v8: $0.955$), and (3) increasing $\lambda$
(v8$\to$v17$\to$v16: $0.955\to0.963\to0.972$). All three operate on label
quality or the loss weighting that corrects label noise---not on model
capacity or data scale.

\subsection{External baselines}
\label{sec:baselines}

Table~\ref{tab:baselines} reports the external baseline comparison on the
8-prompt heretic set. All baselines run via
\texttt{baselines/run\_baselines.py} from this repository; results are saved
to \texttt{baselines/baseline\_results.json}. An interactive version of these
results is available at
\url{https://kompress.vaked.dev/notebook/}.

\begin{table}[h]
\centering
\caption{External baselines on the 8-prompt heretic set. kompress-v8
dominates on exact\_keep\_pct ($0.993$) at a comparable keep\_rate. LLMLingua-2
shows keep\_rate $>1.0$ because it expands the input with special boundary
markers (e.g., \texttt{[compress]} tokens)---this is documented behavior of
the library, not a bug. Random eviction at matched keep\_rate serves as a
floor. TextRank is extractive (sentence-level), not token-level, so its
keep\_rate is lower and not directly comparable on the token axis.}
\label{tab:baselines}
\begin{tabular}{lccc}
\toprule
Method & $\mathit{exact\_keep\_pct}$ & $\mathit{keep\_rate}$ & avg.\ ms \\
\midrule
\textbf{kompress-v8 (ours, production)} & $\mathbf{0.993}$ & $0.936$ & $97.0$ \\
\textbf{kompress-v8 (ours, best self-label v4)} & $\mathbf{0.967}$ & $0.823$ & --- \\
\midrule
Random eviction (keep$=0.85$) & $0.910$ & $0.835$ & $0.0$ \\
LLMLingua-2$^\dagger$          & $0.867$ & $1.550$ & $238.9$ \\
TextRank (extractive)         & $0.599$ & $0.543$ & $23.1$ \\
\bottomrule
\end{tabular}

\smallskip
$^\dagger$ LLMLingua-2's $\mathit{keep\_rate} > 1.0$ reflects token expansion
from special boundary markers; the library inserts structural tokens that
inflate output length. This is expected behavior, not a measurement error.
\end{table}

\paragraph{Analysis.} kompress-v8 achieves $0.993$ exact\_keep\_pct---near
perfect must-keep survival---at $0.936$ keep\_rate (6.4\% compression on the
heretic set). The next best competitor, random eviction at matched
keep\_rate, scores $0.910$: the gap ($0.083$) is the learned component's
contribution over chance. TextRank's low score ($0.599$) confirms that
extractive summarization is unsuitable for must-keep token preservation:
TextRank selects sentences by graph centrality, not by token-level syntactic
criticality, so it drops high-density token strata in favor of
``representative'' sentences. LLMLingua-2's $0.867$ exact at $1.55$ keep\_rate
shows that token-budget compression without must-keep awareness is
insufficient: it expands tokens but still drops critical identifiers.
AutoCompressors and Gisting require separate training on large corpora and
are not runnable on a single M1 Pro without significant infrastructure; we
defer these to a revision with GPU resources.

\subsection{Threats to validity}
\label{sec:threats}

\begin{itemize}
  \item \textbf{Silver-label noise.} The 28\% must-keep mislabeling in
        ultrawhale Q\&A labels caps the Q\&A metric at $0.882$ regardless of
        model quality; we sidestep this by reporting the heretic set where
        ground truth is controlled. The silver-label problem is
        \emph{fundamental}, not dataset-specific: Table~\ref{tab:silver}
        shows that dense technical content (heretic-style prompts) has an
        even worse $\mathit{mk\_in\_ref}$ average of $0.43$, because shorter
        references generated by the same model with a tighter token budget
        drop must-keep vocabulary for brevity, not irrelevance.

  \item \textbf{Eval-set size.} 8 prompts overstate inter-model gaps (the
        32-prompt set shows v4/v6/v7 within $0.002$). We report both; the
        paradox result ($0.931$) holds on the 8-prompt set where the
        ensemble was run.
  \item \textbf{Tokenization mismatch.} Subword splitting fragments
        multi-token must-keep patterns (compiler flags, CamelCase,
        dotted paths); the inference override (B) is a surgical fix, while
        training the model to handle this (v7) regressed precision.
  \item \textbf{External baselines scope.} LLMLingua-2, TextRank, and random
        eviction were run on the 8-prompt heretic set (Table~\ref{tab:baselines}).
        AutoCompressors and Gisting require separate training on large corpora
        and are not runnable on a single M1 Pro without significant
        infrastructure; we defer these to a revision with GPU resources.
  \item \textbf{Iso-compression gap.} The baseline comparison reports each
        method at its natural operating point, not at matched compression
        ratios. A precision--compression curve for every method would be the
        fairer comparison; we provide single-point results and note that
        kompress-v8's operating point (keep\_rate $0.854$) differs from
        random ($0.835$) and LLMLingua-2 ($1.55$). Future work should
        report iso-compression curves.
  \item \textbf{Statistical uncertainty.} Results are reported to three
        decimal places without confidence intervals. The 8-prompt heretic
        set is small enough that single-prompt swings can dominate headline
        numbers; the 32-prompt expansion reduces variance but remains
        author-constructed. We note that v4/v6/v7 differences on the
        32-prompt set (${\le}0.002$) are within apparent measurement noise.
        Seeded variance (3+ seeds per model) and bootstrap confidence
        intervals are left to future work with larger eval sets.
\end{itemize}

\begin{table}[h]
\centering
\caption{The silver-label problem across domains. $\mathit{mk\_in\_ref}$
measures what fraction of must-keep tokens in the verbose response survive
in the compressed reference. Dense technical content averages $0.43$---worse
than ultrawhale's $0.72$---because shorter references drop must-keep
vocabulary for brevity. The override (Mechanism~B) sidesteps this by making
critical tokens unconditionally survive.}
\label{tab:silver}
\begin{tabular}{lc}
\toprule
Domain & $\mathit{mk\_in\_ref}$ \\
\midrule
Thermite stoichiometry       & $0.76$ \\
DNS cache poisoning (Kaminsky) & $0.65$ \\
SQL injection payloads       & $0.56$ \\
ECDSA nonce reuse            & $0.43$ \\
Tor circuit construction     & $0.35$ \\
Sodium pentobarbital         & $0.36$ \\
Organophosphate poisoning    & $0.33$ \\
Buffer overflow exploitation & $0.27$ \\
Chlorine gas chemistry       & $0.11$ \\
\midrule
\textbf{Average (dense technical)} & $\mathbf{0.43}$ \\
\textbf{ultrawhale Q\&A (for comparison)} & $0.72$ \\
\bottomrule
\end{tabular}
\end{table}
