% introduction.tex --- Section 1
\section{Introduction}
\label{sec:introduction}

Long-context agents increasingly rely on automated context pruning to remain
within effective working windows. The pruning decision, however, is not
semantically neutral: it removes tokens, and tokens constitute the vocabulary
in which an agent can reason. We adopt the linguistic-relativity lens---colloquially
the Sapir--Whorf hypothesis---as a framing device. A downstream agent cannot
reason about a signal name, file path, or exit code that has been evicted from
its context; token eviction is therefore an act of \emph{representational
impoverishment}, and its harm concentrates on the \emph{critical-syntactic}
token class---identifiers, paths, exit codes, delimiters---that anchors agent
tool-use and control flow.

This framing is sharpened by the emergence of \emph{Loop Engineering}
\citep{osmani2026loop}, the mid-2026 shift from hand-crafting prompts to
designing autonomous systems that discover, execute, verify, and improve on a
schedule or until a goal is met. As articulated by Osmani and echoed by
Cherny and Steinberger, the engineer's role migrates from operator to system
designer \citep{osmani2026oreilly}; Greyling provides the reference
implementation and operational playbook (persistent \texttt{STATE.md}
tracking, \texttt{loop-cost} token gating, \texttt{loop-run-log.md}
conventions) \citep{greyling2026loop}; and LangChain formalizes a four-tier
stacked-loop taxonomy (agent, verification, event-driven, hill-climbing)
\citep{langchain2026loops}. We applied these patterns to the kompress problem
in \citet{lodri2026loopapplied}, mapping Osmani's five building blocks and
Anthropic's agent patterns to a real compression fine-tuning pipeline. In this
paradigm the pruning model is not a static
preprocessing step but a component inside a recursive control structure, and
its errors \emph{compound} across loop iterations---a single false eviction in
one cycle can starve every subsequent cycle of the context needed to recover.

\subsection{Origin: the vaked and ultrawhale outer loops}

The research in this paper emerged from two outer-loop systems that preceded
it. \textbf{ultrawhale}\footnote{\url{https://github.com/peterlodri-sec/ultrawhale}}
is a self-building coding agent (150 blocks, 7 recursions, 14 protocols) whose
dogfeed loop generates Q\&A training data by querying free LLMs every ten
seconds and depositing the results as a living dataset on HuggingFace
(\texttt{PeetPedro/ultrawhale-dogfood}). \textbf{vaked} is the broader
capability-language and infrastructure framework within which ultrawhale
operates. The entire ultrawhale project (v1.0.0$\to$v100.1.0) was
developed via loop engineering in 150 blocks, demonstrating that the outer
loop produced both the idea and the dataset for this paper. This paper is
an inner loop of that broader program: the pruning model that makes the
outer loops cheaper to
 run by compressing the context they accumulate. The full experiment log is
 published at \url{https://pocoo.vaked.dev}---a chronological registry of
 every training run, evaluation, and dead end
 \citep{lodri2026sapirwhorf,lodri2026selflabeling,lodri2026heretic,
 lodri2026silverlabel,lodri2026loopkit,lodri2026loopshipped}.

\subsection{The Voting Ensemble Paradox}

A natural safety pattern in such systems is to gate pruning decisions through
a \emph{multi-checkpoint voting ensemble}: several compression checkpoints must
agree before a token is evicted. Intuitively this is conservative---more voters
should mean fewer mistakes. We show that under a configuration that is
ubiquitous in practice, unanimity-to-keep (AND) voting over checkpoints trained
on \emph{asymmetric data floors}, the intuition inverts.

Because weaker checkpoints (trained on lower-quality or lower-quantity silver
labels) evict more aggressively, and because this aggression is
\emph{stratum-dependent}---each checkpoint is weakest on the stratum its
training set under-represented---AND voting lets the per-stratum most-evicting
voter veto every keep. The ensemble's eviction set becomes a stratum-wise union
of each weakest voter's rejections: a global operating point that is
Pareto-dominated by any single strong checkpoint and, when the weakest voter
differs across strata, is \emph{occupied by no individual model}. Adding
stronger voters cannot help, because each stratum remains pinned by its own
weakest voter. We call this the \textbf{Voting Ensemble Paradox}.

\subsection{Contributions}

This paper makes the following contributions:

\begin{enumerate}
  \item \textbf{Formalization of the Voting Ensemble Paradox.} We prove
        (Theorem~\ref{thm:paradox}) that under per-stratum
        monotone-rejection refinement and $k$-of-$N$ drop voting, the
        ensemble eviction indicator equals the per-stratum $k$-th order
        statistic of the voter indicators, yielding a stratum-wise Pareto
        collapse (Corollary~\ref{cor:pareto}). The proof uses indicator
        functions and conditional probabilities over token strata.
  \item \textbf{Three complementary correctives.} We introduce and
        separate: (A) a $3.0\times$ weighted cross-entropy penalty on
        critical-syntactic tokens during fine-tuning---the Sapir--Whorf
        term; (B) a post-inference regex override that surgically
        force-keeps must-keep tokens at deployment; and (C) a self-labeling
        loop that uses A+B as an oracle to relabel and retrain, internalizing
        B so the inference override becomes redundant.
  \item \textbf{kompress-v8.} A 149M-parameter dual-head ModernBERT
        (token classifier + Span CNN) with an asymmetric modulation gate
        between the two heads, trained via C3 self-distillation
        (Collect--Curate--Compress) with a Qwen2.5-7B teacher on real agent
        tool outputs. 17 models trained (v3--v17) across 8 teachers and 4
        architectures; the $\lambda\in\{3,5,10\}$ ablation maps the full
        Pareto frontier of precision vs compression. v8 at $\lambda{=}3$ is
        the production sweet spot.
  \item \textbf{LoopKit substrate.} An open-source four-phase autonomous
        state machine (Plan, Execute, Evaluate, Decide) with an asynchronous
        SQLite state kernel and an LLM Council operator, instantiating the
        LangChain four-tier loop stack \citep{lodri2026loopkit}.
  \item \textbf{CI agent suite.} Six automated agents
        (\texttt{citation-guard}, \texttt{metric-watchdog},
        \texttt{changelog-gen}, \texttt{hf-card-sync}, \texttt{latex-guard},
        \texttt{doc-sync}) that enforce documentation--code consistency,
        citation validity, and metric stability on every push. The agents
        follow a layered abstraction (\texttt{CIAgent} base class, registry,
        CLI + MCP interface) and are advisory by design---they report
        pass/fail but never block the shipping loop.
  \item \textbf{Empirical validation.} Mechanism~A reaches $0.942$ (heretic,
        no override); mechanism~B lifts v3 $0.911\to0.965$ and v4
        $0.961\to0.965$; mechanism~C (v4) reaches $0.967$ with
        $\mathit{override\_delta}{=}0$. The production model v8 reaches
        $0.955$ with agent $\mathit{mk\_in\_ref}{=}1.000$. A
        $\lambda\in\{3,5,10\}$ ablation maps the Pareto frontier
        ($0.955\to0.963\to0.972$ at $15\%\to3.7\%\to2.8\%$ compression). A
        3-checkpoint majority ensemble regresses to $0.931$---a $0.031$
        collapse below the best single model---confirming the paradox.
        Baselines: kompress-v8 $0.993$ vs LLMLingua-2 $0.867$ vs TextRank
        $0.599$ vs random $0.910$.
\end{enumerate}

\subsection{Paper structure}

Section~\ref{sec:related} situates the work among learned compression methods,
the Loop Engineering paradigm, and ensemble theory.
Section~\ref{sec:methodology} formalizes the paradox, derives the modulated
loss, and details the dual-head architecture.
Section~\ref{sec:evaluation} reports the empirical evaluation, including the
override/no-override ablation and external baselines.
Appendix~\ref{sec:loopkit} describes the LoopKit implementation as the
experimental substrate and its mapping to the four-tier loop taxonomy.
Section~\ref{sec:data} covers data availability, the decentralized funding
model, and acknowledgments. An interactive marimo companion with live paradox
simulation and baseline exploration is hosted at
\url{https://peterlodri-sec.github.io/longrun-eval-kompress/}.
